---
title: "Vader Sentiment Analysis"
format: html
---

## VADER Sentiment Analysis

We use VADER to perform sentiment analysis on all tweets.

VADER gives us a negativity score, positivity score, neutral score, and a compound score which is a combination of the three. Positivity score, negativity score, and neutral score all range from 0 to 1. Compound score ranges from -1 to 1, with 1 being most positive and -1 being most negative.

```{r Setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = T, cache = T, fig.width=8, fig.height=4,
                      warning = F)
options(scipen = 0, digits = 3)  ## controls base R output

## Package setup
if(!require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, vader)
```

We analyze Biden first.

```{r}
biden2020 <- read_csv("data/biden2020.csv", show_col_types = FALSE)  

vscores <- biden2020$text %>% lapply(get_vader)

biden2020sentiment <- biden2020 %>% mutate(
  compound = vscores %>% sapply(function(v) { as.numeric(v["compound"]) }),
  pos = vscores %>% sapply(function(v) { as.numeric(v["pos"]) }),
  neu = vscores %>% sapply(function(v) { as.numeric(v["neu"]) }),
  neg = vscores %>% sapply(function(v) { as.numeric(v["neg"]) }),
)

biden2020sentiment

write.CSV(biden2020sentiment, file="data/biden2020sentiment.csv")
```

Then, we analyze Trump.

```{r}
trump2020 <- read_csv("data/trump2020.csv", show_col_types = FALSE)  

tscores <- trump2020$text %>% lapply(get_vader)

trump2020sentiment <- trump2020 %>% mutate(
  compound = tscores %>% sapply(function(v) { as.numeric(v["compound"]) }),
  pos = tscores %>% sapply(function(v) { as.numeric(v["pos"]) }),
  neu = tscores %>% sapply(function(v) { as.numeric(v["neu"]) }),
  neg = tscores %>% sapply(function(v) { as.numeric(v["neg"]) }),
)

trump2020sentiment

write.csv(trump2020sentiment, file="data/trump2020sentiment.csv")
```
